---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r}
DT<-read.csv("C:/Users/Romain/Desktop/datasciencecoursera/Reproductible Research/Project 1/activity.csv")
head(DT)
DT$date<-as.Date(DT$date,format = "%Y-%m-%d")
class(DT$date)

DT<-DT[complete.cases(DT),]
```

## What is mean total number of steps taken per day?
```{r}
sumsteps<-tapply(DT$steps,DT$date,sum)

sumsteps<-as.data.frame.table(sumsteps)

hist(sumsteps$Freq,breaks = 10,col="red",
     xlim = c(41,21194),
     xlab = "Total Steps by day",
     main = "Total number of steps taken per day")

abline(v=median(sumsteps$Freq),col="blue")
median(sumsteps$Freq)
abline(v=mean(sumsteps$Freq),col="black")
mean(sumsteps$Freq)

```

## What is the average daily activity pattern?
```{r}
stepbyinterval<-tapply(DT$steps,DT$interval,mean)

stepbyinterval<-as.data.frame.table(stepbyinterval)

plot(stepbyinterval$Var1,stepbyinterval$Freq, "l",
     xlab="5-minute interval",
     ylab="average number of steps",
     main="The average daily activity pattern")

lines(as.numeric(stepbyinterval$Var1),stepbyinterval$Freq, "l", col = "black")

stepbyinterval[which(stepbyinterval$Freq==max(stepbyinterval$Freq)),1]

abline(v=stepbyinterval[which(stepbyinterval$Freq==max(stepbyinterval$Freq)),1],
       col="red")
```

## Imputing missing values
The strategy chosen for imputing missing data is to fill by the mean of each interval
corresponding.
```{r}
DT<-read.csv("C:/Users/Romain/Desktop/datasciencecoursera/Reproductible Research/Project 1/activity.csv")

DTna<-DT[is.na(DT$steps), ]

newDT<-data.frame(steps=stepbyinterval
                  [match(DTna$interval,stepbyinterval$Var1),2],
                  date=DTna$date,
                  interval=DTna$interval)

DTcc<-DT[complete.cases(DT),]

fDT<-rbind(newDT,DTcc)

fDT$date<-as.Date(fDT$date,format = "%Y-%m-%d")

sumsteps2<-tapply(fDT$steps,fDT$date,sum)
sumsteps2<-as.data.frame.table(sumsteps2)

hist(sumsteps2$Freq,breaks = 10,col="red",
     xlim = c(41,21194),xlab = "Total Steps by day",
     main = "Total number of steps taken per day")

abline(v=median(sumsteps2$Freq),col="blue")
abline(v=mean(sumsteps2$Freq),col="black")
```
This strategy doesn't doesn't change the Histogram.

## Are there differences in activity patterns between weekdays and weekends?
First we create a new column with the factor : weekdays and weekends.
```{r}
library(dplyr)

wkDT<-mutate(fDT,day=weekdays(fDT$date))

NewVar <- factor(rep(NA, length(wkDT$day) ), 
                 levels=c("weekday", "weekend") )   
NewVar[ wkDT$day %in% c("lundi", "mardi", "mercredi","jeudi","vendredi")]<-"weekday"
NewVar[ is.na(NewVar) ]<-"weekend"

wkDT<-cbind(wkDT,NewVar)

wkDT1<-subset(wkDT,wkDT$NewVar=="weekday")
wkDT1interval<-tapply(wkDT1$steps,wkDT1$interval,mean)
wkDT1interval<-as.data.frame.table(wkDT1interval)

wkDT2<-subset(wkDT,wkDT$NewVar=="weekend")
wkDT2interval<-tapply(wkDT2$steps,wkDT2$interval,mean)
wkDT2interval<-as.data.frame.table(wkDT2interval)

plot(wkDT2interval$Var1,wkDT2interval$Freq, "l",xlab="5-minute interval",
     ylab="Number of steps",main="weekend")
lines(as.numeric(wkDT2interval$Var1),wkDT2interval$Freq, "l", col = "black")

plot(wkDT1interval$Var1,wkDT1interval$Freq, "l",xlab="5-minute interval",
     ylab="Number of steps",main="weekday")
lines(as.numeric(wkDT1interval$Var1),wkDT1interval$Freq, "l", col = "black")

```
We can see that the patterns are different